---
title: "Final Presentation"
author: "Chris Kardish and Marie Agosta"
date: "November 18, 2016"
output: pdf_document
---

This is where we clean and prepare the data

```{r}

#rm(list = ls()) this command clears your environment
#setwd('C:/Users/Chris/Documents/GitHub/final_project')
#setwd('C:/Users/Chris/Documents/R/col_soc_science')

source("final_cleaning.R")

# calling appropriate packages 

library(foreign) 
library(survey)
library(rockchalk)

# R will crash if a primary sampling unit (psu) has a single observation
# so we set R to produce conservative standard errors instead of crashing
options( survey.lonely.psu = "adjust" )

#import data from the ANES
anes <- read.dta("anes_timeseries_2012_Stata12.dta")

# to delete ages that do not fit into any of our generations, we delete the unnecesary ages
#table(anes$dem_age_r_x) #age
anes <- anes[!(anes$dem_age_r_x <= -2 | anes$dem_age_r_x >= 88),] 

#We renamed the levels pertaining to if someone voted and then we coded them to be binary, keeping only Yes/no, changes the variable to a numeric dummy, with 1 as "Yes"
#table(anes$postvote_rvote)
levels(anes$postvote_rvote) <- c("Refused", "Don't know", "Incomplete", "Nonresponsive", "Missing", "Inapplicable", "No", "No", "No", "Yes")
anes <- anes[(anes$postvote_rvote == "Yes" | anes$postvote_rvote == "No"),] 
anes$postvote_rvote <-  droplevels(anes$postvote_rvote)
anes$postvote_rvote <- as.numeric(anes$postvote_rvote) 
anes$postvote_rvote[anes$postvote_rvote == 1] <- 0
anes$postvote_rvote[anes$postvote_rvote == 2] <- 1

#We renamed the levels pertaining to education and recoded them as numeric ranging from 
#1-5 (1 being below high school) and 5 (graduate education)
#table(anes$dem_edugroup_x) #education
levels(anes$dem_edugroup_x) <- c("Refused", "Don't know", "Data missing", "Below high school", "High school", "Some post-high", "Bachelor", "Graduate") 
anes <- anes[(anes$dem_edugroup_x == "Below high school" | 
                anes$dem_edugroup_x == "High school" | 
                anes$dem_edugroup_x == "Some post-high" | 
                anes$dem_edugroup_x == "Bachelor" |
                anes$dem_edugroup_x == "Graduate"),] #keeps only these groups
anes$dem_edugroup_x <-  droplevels(anes$dem_edugroup_x)
anes$dem_edugroup_x <- as.numeric(anes$dem_edugroup_x)

#We renamed the labels pertaining to race, removed all missing data, and created dummies of black
#and hispanic
#table(anes$dem_raceeth_x) #race
levels(anes$dem_raceeth_x) <- c("Data missing", "White", "Black", "Asian" , "Native American or Alaska Native", "Hispanic","Other") 
anes <- anes[!(anes$dem_raceeth_x == "Data missing"),] 
anes$black <- anes$dem_raceeth_x == "Black"
anes$black <- as.numeric(anes$black)
anes$hispanic <- anes$dem_raceeth_x == "Hispanic"
anes$hispanic <- as.numeric(anes$hispanic)


#We renamed the labels pertaining to voting in the 2008 election, dropped NA ones, and recoded #the remaining yes and no to be 1 or 0. 
#table(anes$interest_voted2008) #voting in past election (2008)
levels(anes$interest_voted2008) <- c("Refused", "Don't know", "Yes", "No")
anes <- anes[(anes$interest_voted2008 == "Yes" | anes$interest_voted2008 == "No"),] 
anes$interest_voted2008 <-  droplevels(anes$interest_voted2008)
anes$interest_voted2008 <- as.numeric(anes$interest_voted2008) 
anes$interest_voted2008[anes$interest_voted2008 == 2] <- 0

#We created a dummy variable for gender, with 1 as "female"
#table(anes$gender_respondent_x) #gender
anes$female <- anes$gender_respondent_x == "2. Female" #creates dummy gender #variable
anes$female <- as.numeric(anes$female) 

#table(anes$inc_incgroup_pre) #income groups
anes <- anes[!(anes$inc_incgroup_pre == "-9. Refused"),] 
anes <- anes[!(anes$inc_incgroup_pre == "-8. Don't know"),] 
anes <- anes[!(anes$inc_incgroup_pre == "-2. Missing; IWR mistakenly entered '2' in place of DK code for total income"),] 
anes$inc_incgroup_pre <- combineLevels(anes$inc_incgroup_pre, c("01. Under $5,000", "02. $5,000-$9,999", "03. $10,000-$12,499", "04. $12,500-$14,999", "05. $15,000-$17,499", "06. $17,500-$19,999"), "Under 20k")
anes$inc_incgroup_pre <- combineLevels(anes$inc_incgroup_pre, c("07. $20,000-$22,499", "08. $22,500-$24,999", "09. $25,000-$27,499", "10. $27,500-$29,999", "11. $30,000-$34,999", "12. $35,000-$39,999", "13. $40,000-$44,999"), "20k-45k")
anes$inc_incgroup_pre <- combineLevels(anes$inc_incgroup_pre, c("14. $45,000-$49,999", "15. $50,000-$54,999", "16. $55,000-$59,999", "17. $60,000-$64,999"), "45k-65k")
anes$inc_incgroup_pre <- combineLevels(anes$inc_incgroup_pre, c("18. $65,000-$69,999", "19. $70,000-$74,999", "20. $75,000-$79,999", "21. $80,000-$89,999"), "65k-90k")
anes$inc_incgroup_pre <- combineLevels(anes$inc_incgroup_pre, c("22. $90,000-$99,999", "23. $100,000-$109,999", "24. $110,000-$124,999"), "90k-125k")
anes$inc_incgroup_pre <- combineLevels(anes$inc_incgroup_pre, c("25. $125,000-$149,999", "26. $150,000-$174,999", "27. $175,000-$249,999"), "125-250k")
levels(anes$inc_incgroup_pre) <- c("Under 20k", "20k-45k", "45k-65k", "65k-90k", "90k-125k", "125-250k", "28. $250,000 or more")
anes$income <- as.numeric(anes$inc_incgroup_pre)


#To simplify our task we create a smaller data set that includes all relevant variables
#, which are listed below the code. 

anes_small <- data.frame(anes$caseid, anes$dem_age_r_x, anes$dem_edugroup_x, anes$black, anes$hispanic, anes$income, anes$female, anes$interest_voted2008, anes$postvote_rvote, anes$weight_full, anes$psu_full, anes$strata_full)

# We named the variables so they are very easy to understand 
colnames(anes_small) <- c("caseID", "age", "education", "black","hispanic", "income", "female", "vote_2008", "vote_2012", "weights", "psu", "strata" )

```

Creating the age-band subsets from the smaller ANES data set that contains only 
variables of interest.

```{r}

anes_genY <- subset(anes_small, anes$dem_age_r_x > 17 & anes$dem_age_r_x < 33) 
#creates Millennial subset 
anes_genX <- subset(anes_small, anes$dem_age_r_x > 32 & anes$dem_age_r_x < 48) 
#creates Generation X subset 
anes_boomer <- subset(anes_small, anes$dem_age_r_x > 47 & anes$dem_age_r_x < 67) 
#creates Baby Boomer subset 
anes_silent <- subset(anes_small, anes$dem_age_r_x > 66 & anes$dem_age_r_x < 88) 
#creates Silent Generation subset

#set up an objects to run with the logit models that allow for results to be 
#statistically representative of the population through weighting 

ANESdesign_genY <- svydesign(~psu ,  strata = ~strata , data = anes_genY , weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_genX <- svydesign(~psu ,  strata = ~strata , data = anes_genX , weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_boomer <- svydesign(~psu ,  strata = ~strata , data = anes_boomer, weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_silent <- svydesign(~psu ,  strata = ~strata , data = anes_silent, weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_overall <- svydesign(~psu ,  strata = ~strata , data = anes_small, weights = ~weights, variables = NULL, nest = TRUE)
```

Running logistic regression model with confidence intervals

```{r}

#Running a logit, generation by generation. Then looking at a summary of the results, follows by looking at and creating an object for the coeffs of the results. Then converting those results to probabilities by exp(ing) them 

#Millennials
M_genY <- svyglm(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, design = ANESdesign_genY, family = "quasibinomial") 
summary(M_genY)
genY_coef <- coef(M_genY)
exp(genY_coef)

#Gen X
M_genX <- svyglm(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, design = ANESdesign_genX, family = "quasibinomial") 
summary(M_genX)
genX_coef <- coef(M_genX)
exp(genX_coef)

#Baby Boomers
M_boomer <- svyglm(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, design = ANESdesign_boomer, family = "quasibinomial") 
summary(M_boomer)
boomer_coef <- coef(M_boomer)
exp(boomer_coef)

#Silent
M_silent <- svyglm(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, design = ANESdesign_silent, family = "quasibinomial") 
summary(M_silent)
silent_coef <- coef(M_silent)
exp(silent_coef)

#overall data set
M <- svyglm(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, design = ANESdesign_overall, family = "quasibinomial") 
summary(M)


newdata1 <- with(anes_small, data.frame(education = 1:5, black = 0, hispanic = 0, income = 4, female = 0, vote_2008 = 1))

  
  
#Looking at confidence intervals of parameter point estimates 

confint(M_genY)
confint(M_boomer)
confint(M_genX)

#Giving odds ratios with confidence intervals in a table 

exp(cbind(OR = coef(M_genY), confint(M_genY)))

#For presentation... because stargazer doesn't output to presenation 

Millennials <- exp(cbind(OR = coef(M_genY), confint(M_genY)))
Mill <- data.frame(Millennials)
colnames(Mill) <- c("Predicted Probabilites", "Lower Bound", "Upper Bound") 
library(knitr)

kable(Mill)
kable(Mill, align = 'c', digits = 2, caption = 'Predicted Probabilities for Millennials')

```

Code for stargazer table comparing odds-ratios of coefficients across models. 
Not finished with it yet, though. 
```{r}
library(stargazer)

#to look at the stargazer, try the one below, which uses "text." The one below only appears
#when you knit to PDF.

stargazer(M_genY, M_genX, M_boomer, M_silent, apply.coef = exp, type = 'text',
          dep.var.labels = "Probability of voting in 2012", 
          column.labels = c("Gen Y", "Gen X", "Boomers", "Silent"))

stargazer(M_genY, M_genX, M_boomer, M_silent, apply.coef = exp,
          dep.var.labels = "Probability of voting in 2012", 
          column.labels = c("Gen Y", "Gen X", "Boomers", "Silent"))

```

Attempt at the simulation stuff he's into
```{r}
library(MASS)
library(dplyr)
library(Zelig)
#Steps. 1) run model but using Zelig, which is able to use survey weights
#2) use "setx" function to set the values for the simulation
#3) run the simulation using the parameters you specified (sim)
#4) plot using ci.plot

#this is black millennial female at various levels of education and moderate income

Z_genY <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, data = anes_genY)
setZ_genY_black <- setx(Z_genY, education = 1:5, black = 1, hispanic = 0, income = 3, 
                        female = 1, vote_2008 = 0)
simZ_genY_black <- sim(Z_genY, x = setZ_genY_black)
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", xlab = "Education Level", 
        main = "Black Millennial Woman", legcol = "black", leg = 1,) 

#With Labels (that don't work... ) 
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", xlab = "Education Level", axis(side = 1, at = 1:5, tick = TRUE, labels = c("Below HS", "HS", "Some post-HS", "B", "Masters", )

                                                                         
#same thing but a black baby boomer

Z_boomer <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, data = anes_boomer)
<<<<<<< HEAD
setZ_boomer_black <- setx(Z_genY, education = 1:5, black = 1, hispanic = 0, income = 3, 
                          female = 1, vote_2008 = 0)
simZ_boomer_black <- sim(Z_boomer, x = setZ_boomer_black)
ci.plot(simZ_boomer_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Baby Boomer Woman")
=======
setZ_boomer_black <- setx(Z_genY, education = 1:5, black = 1, hispanic = 0, income = 3, female = 1, vote_2008 = 0)
simZ_boomer_black <- sim(Z_boomer, x = setZ_boomer_black)
ci.plot(simZ_boomer_black, ylim = 0:1, ylab = "Probability of voting", xlab = "Education Level",
        main = "Black Baby Boomer Woman", legcol = "black", leg = 1,)

#With Labels (that don't work... )
ci.plot(simZ_boomer_black, ylim = 0:1, ylab = "Probability of voting", xlab = "Education Level", axis(side = 1, at = 1:5, tick = TRUE, 
        labels = c("Below HS", "HS", "Some post-HS", "Bacholer", "Masters", )

>>>>>>> origin/master
#CONCLUSION: plots look really damn similar, which may be expected given that 
#the coefficients for Boomers and Millennials do show striking similarities, with #the exception of race stuff. So the fact that a black person model looks so similar #between the two generations is perhaps a bit odd. 

#trying now a white man with a high school diploma across different levels of income

#First, Millennial

setZ_genY_white <- setx(Z_genY, income = 1:7, education = 2, black = 0, hispanic = 0, 
<<<<<<< HEAD
                        female =0, vote_2008 = 1)
=======
                        female =0, vote_2008 = 0)

>>>>>>> origin/master
simZ_genY_white <- sim(Z_genY, x = setZ_genY_white)

ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting",
        main = "White Male Millenial with HS Edu", legcol = "black", leg = 1, )

#With labels 
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", axis(side = 1, at = 1:7, tick = TRUE, labels = c("20k", "20-40k", "45k-65k", "65k-90k", "90k-125k", "125-250k", "250,000k or more", ) 

#Now, Baby Boomer
setZ_boomer_white <- setx(Z_boomer, income = 1:7, education = 2, black = 0, hispanic = 0, 
                        female =0, vote_2008 = 1)
simZ_boomer_white <- sim(Z_boomer, x = setZ_boomer_white)
ci.plot(simZ_boomer_white, ylim = 0:1, ylab = "Probability of voting",
        main = "White Male Babyboomer with HS Edu", legcol = "black", leg = 1, )

#With Labels (that don't work... )
ci.plot(simZ_boomer_white, ylim = 0:1, ylab = "Probability of voting",axis(side = 1, at = 1:7, tick = TRUE, labels = c("20k", "20-40k", "45k-65k", "65k-90k", "90k-125k", "125-250k", "250,000k or more", ) 

#again, plots between Boomers and Millennials look really damn similar

```

Now playing with and plotting predicted probabilities based on the model

```{r}

predicted_fit <- function(alpha, beta1, beta2, beta3, beta4, beta5, beta6, 
    x1, x2, x3, x4, x5, x6, 
    se1, se2, se3, se4, se5, se6)
{
lb_fun <- function(coef, se){
    lb <- coef - (1.96 * se)
    return(lb)
}

ub_fun <- function(coef, se){
    ub <- coef + (1.96 * se)
    return(ub)
}

fit <- alpha + beta1 * x1 + beta2 * x2 + beta3 * x3 + beta4 * x4 + beta5 * x5
lb <- alpha + lb_fun(beta1, se1) * x1  + lb_fun(beta2, se2) * x2  + lb_fun(beta3, se3) * x3
        lb_fun(beta4, se4) * x4 + lb_fun(beta5, se5) * x5 + lb_fun(beta6, se6) * x6
ub <- alpha + ub_fun(beta1, se1) * x1  + ub_fun(beta2, se2) * x2  + ub_fun(beta3, se3) * x3
        ub_fun(beta4, se4) * x4 + ub_fun(beta5, se5) *x5 + ub_fun(beta6, se6) *x6
predictions <- c(lb, fit, ub)

probs <- function(x) {(exp(x)) / (1 + exp(x)) }
predictions <- probs(predictions)
return(predictions) 
}

#let's start by comparing a black Millennial woman and a black Baby Boomer woman
#across different levels of education. Let's assume she didn't vote in 2008 and
#her income is moderate (3). The lines of code below simply gives us our predicted
#probabilities, along with confidence intervals.

black_genY_ed1 <- predicted_fit(-1.42663, 0.27571, 0.10578, 1.08370, 0.97081, 0.63016, 
                                0.19622, 1, 1, 0, 1, 0, 3, 0.13361, 0.23487, 0.24849, 
                                0.39274, 0.35245, 0.07876)

black_genY_ed2 <- predicted_fit(-1.42663, 0.27571, 0.10578, 1.08370, 0.97081, 0.63016, 
                                0.19622, 2, 1, 0, 1, 0, 3, 0.13361, 0.23487, 0.24849, 
                                0.39274, 0.35245, 0.07876)

black_genY_ed3 <- predicted_fit(-1.42663, 0.27571, 0.10578, 1.08370, 0.97081, 0.63016, 
                                0.19622, 3, 1, 0, 1, 0, 3, 0.13361, 0.23487, 0.24849, 
                                0.39274, 0.35245, 0.07876)

black_genY_ed4 <- predicted_fit(-1.42663, 0.27571, 0.10578, 1.08370, 0.97081, 0.63016, 
                                0.19622, 4, 1, 0, 1, 0, 3, 0.13361, 0.23487, 0.24849, 
                                0.39274, 0.35245, 0.07876)

black_genY_ed5 <- predicted_fit(-1.42663, 0.27571, 0.10578, 1.08370, 0.97081, 0.63016, 
                                0.19622, 5, 1, 0, 1, 0, 3, 0.13361, 0.23487, 0.24849, 
                                0.39274, 0.35245, 0.07876)

black_boomer_ed1 <- predicted_fit(-1.49411, 0.32012, 0.18731, 2.67234, -0.73006, -0.12489,
                                  0.17901, 1, 1, 0, 1, 0, 3, 0.12001, 0.25347, 0.25373,
                                  0.35665, 0.39013, 0.06914)

black_boomer_ed2 <- predicted_fit(-1.49411, 0.32012, 0.18731, 2.67234, -0.73006, -0.12489,
                                  0.17901, 2, 1, 0, 1, 0, 3, 0.12001, 0.25347, 0.25373,
                                  0.35665, 0.39013, 0.06914)

black_boomer_ed3 <- predicted_fit(-1.49411, 0.32012, 0.18731, 2.67234, -0.73006, -0.12489,
                                  0.17901, 3, 1, 0, 1, 0, 3, 0.12001, 0.25347, 0.25373,
                                  0.35665, 0.39013, 0.06914)

black_boomer_ed4 <- predicted_fit(-1.49411, 0.32012, 0.18731, 2.67234, -0.73006, -0.12489,
                                  0.17901, 4, 1, 0, 1, 0, 3, 0.12001, 0.25347, 0.25373,
                                  0.35665, 0.39013, 0.06914)

black_boomer_ed5 <- predicted_fit(-1.49411, 0.32012, 0.18731, 2.67234, -0.73006, -0.12489,
                                  0.17901, 5, 1, 0, 1, 0, 3, 0.12001, 0.25347, 0.25373,
                                  0.35665, 0.39013, 0.06914)

#now we take those predictions and put them in a dataframe that we can use for 
#plotting and other functions
library(dplyr)
blackfem_ed_genY <- rbind(black_genY_ed1, black_genY_ed2, black_genY_ed3, black_genY_ed4,
                   black_genY_ed5)  %>% as.data.frame()
blackfem_ed_boomer <- rbind(black_boomer_ed1, black_boomer_ed2, black_boomer_ed3,
                   black_boomer_ed4, black_boomer_ed5) %>% as.data.frame()
blackfem_ed_genY$education <- c("Below high school", "High school", "Some post-high", 
                                "Bachelor", "Graduate")
blackfem_ed_boomer$education <- c("Below high school", "High school", "Some post-high", 
                                "Bachelor", "Graduate")
colnames(blackfem_ed_genY) <- c("lb", "probability", "education")
colnames(blackfem_ed_boomer) <- c("lb", "probability", "education")
rownames(blackfem_ed_genY) <- c("Below high school", "High school", "Some post-high", 
                                "Bachelor", "Graduate")
rownames(blackfem_ed_boomer) <- c("Below high school", "High school", "Some post-high", 
                                "Bachelor", "Graduate")

# want to have lines for the two generation we are comparing 
library(ggplot2)
#ggplot(blackfem_genY, aes(x = education, y = probability, group = 1)) +
#  geom_ribbon(aes(ymin = LL, ymax = UL), alpha = .2) +
#  geom_line(aes(colour = education), size=1)

ggplot(blackfem_ed_genY, aes(x = factor(1:5), y = probability, group = 1)) +
  geom_line() + 
              geom_line(data = blackfem_ed_boomer, aes(y = probability, color = probability)) +
  labs(x = "Education", y = "Probability of voting", title = "Millennial vs Boomer") +    scale_x_discrete(labels = c("Below high school", "High school", "Some post-high", 
                                "Bachelor", "Graduate")) +
  geom_ribbon(data=blackfem_ed_genY,aes(ymin=lb,ymax=ub),alpha=0.1) +
  geom_ribbon(data=blackfem_ed_boomer,aes(ymin=lb,ymax=ub),alpha=0.3) 

#the code above works. still need to adjust colors and aesthetics. and also to figure out
#why our confidence bands are so huge. makes things look really imprecise. 

```

Assessing the model fit. I draw on the UCLA page here for this: 
http://www.ats.ucla.edu/stat/r/dae/logit.htm.

The test below decides whether the model is a better predictor than a model with just 
an intercept and nothing else (a null model). The survey-weighted one, simply, isn't. It's awful, according to this test. But running our model without the survey weights it's incredible. 

```{r}

with(M_genY, null.deviance - deviance) #gives chi-square value
with(M_genY, df.null - df.residual) #gives degrees of freedom
with(M_genY, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE)) #gives p value

with(M_boomer, null.deviance - deviance) #gives chi-square value
with(M_boomer, df.null - df.residual) #gives degrees of freed
with(M_boomer, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE)) 

#trying the model without survey weights
M_genY2 <- glm(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, data = anes_genY, family = "binomial") 

with(M_genY2, null.deviance - deviance) #gives chi-square value
with(M_genY2, df.null - df.residual) #gives degrees of freed
with(M_genY2, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE)) 

#Hosmer-Lemeshow Goodness of Fit
install.packages("ResourceSelection")
library(ResourceSelection)
hoslem.test(anes_genY$vote_2012, fitted(M_genY))



````


Notes from class Nov. 18

Should have correlation matrix like on slide 22 of the Nov. 18 class

Run the simulated stuff he did for presentation purposes. Need a regression table
and other stuff, but for highlighting the main effects that's simply not
a good idea. Simluation helps show the results and how confident we are
about those results. Gives an expectation of what we would see in the "real world"
based on the assumptions of our model. 
